---
title: "COVID-19 Policy-Compliance in Europe"
author: "Philipp Holz"
#date: "`r format(Sys.time(), '%d %B, %Y')`"
date: "05 January 2021"
output: 
  pdf_document:
    toc: true
    latex_engine: xelatex
    fig_width: 6
    fig_height: 4
bibliography: References.json

---

```{r, echo=FALSE, message = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = FALSE, dev = "cairo_pdf")
```


```{r libraries, include=FALSE}
# Loading the libraries ----
library(dplyr)
library(ggplot2)
library(scales)
library(lme4)
library(broom.mixed)
library(plotly)
library(reshape2)
library(utils)
library(texreg)
library(stargazer)
library(lmtest)
library(extrafont)
loadfonts()
```


```{r data, eval=TRUE, include=FALSE}
# Data Wrangling -----
## Adding the Policy Data -----
### Read in the Data
policy_data <- read.csv("PolicyData.csv")

### convert Date column to date format
policy_data$Date = as.Date(as.character(policy_data$Date), format = "%Y%m%d")

### create dataframe with only national level data
policy_data_countries <- policy_data %>% filter(Jurisdiction == "NAT_TOTAL")

### create dataframe with only national level of Germany, France, UK, Spain, Italy
policy_data_Europe <- policy_data_countries %>% filter(CountryCode == "DEU" |
                                                         CountryCode == "ITA" |
                                                         CountryCode == "FRA" |
                                                         CountryCode == "ESP" |
                                                         CountryCode == "GBR" |
                                                         CountryCode == "NLD" |
                                                         CountryCode == "BEL" |
                                                         CountryCode == "POL" |
                                                         CountryCode == "SWE" |
                                                         CountryCode == "AUT" |
                                                         CountryCode == "IRL" |
                                                         CountryCode == "DNK")

### Only selecting the columns that are interesting
policy_data_Europe_small <- policy_data_Europe %>% select(Date, CountryName, StringencyIndex)


## Adding the Mobility Data ----
### Reading in the data
mobility_data <- read.csv("MobilityData.csv")

### Add a column with an average of the change in sectors, excluding residential and parks

mobility_data <- mobility_data %>% mutate(avg_change = rowMeans(.[,c(6, 7, 9, 10)]))

### force recognition of the date format
mobility_data$date = as.Date(mobility_data$date, format = "%Y-%m-%d")

### Filter so that only country-level data remains
mobility_data_countries <- mobility_data %>% filter(sub.region.1 == "Total" & sub.region.2 == "Total")

### Create mobility data frame for Europe
mobility_data_Europe <- mobility_data_countries %>% filter(country == "Germany" |
                                                             country == "Italy" |
                                                             country == "France" |
                                                             country == "United Kingdom" |
                                                             country == "Spain" |
                                                             country == "Netherlands" |
                                                             country == "Belgium" |
                                                             country == "Poland" |
                                                             country == "Sweden" |
                                                             country == "Austria" |
                                                             country == "Denmark" |
                                                             country == "Ireland")

### renaming the date columns to match
names(mobility_data_Europe)[names(mobility_data_Europe)=="date"] <- "Date"
names(mobility_data_Europe)[names(mobility_data_Europe)=="country"] <- "CountryName"
names(mobility_data_Europe)[names(mobility_data_Europe)=="avg_change"] <- "MobilityIndex"

### Only selecting relevant data
mobility_data_Europe_small <- mobility_data_Europe %>% select(Date, CountryName, MobilityIndex)


## Adding the Cases Data ----

###read the Dataset sheet in
cases_data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")

### renaming the date and countries columns to match
names(cases_data)[names(cases_data)=="dateRep"] <- "Date"
names(cases_data)[names(cases_data)=="countriesAndTerritories"] <- "CountryName"

### Forcing Date format
cases_data$Date = as.Date(as.character(cases_data$Date), format = "%d/%m/%Y")

### create dataframe with only national level of Germany, France, UK, Spain, Italy
cases_data_Europe <- cases_data %>% filter(CountryName == "Germany" |
                                             CountryName == "Italy" |
                                             CountryName == "France" |
                                             CountryName == "Spain" |
                                             CountryName == "United_Kingdom" |
                                             CountryName == "Netherlands" |
                                             CountryName == "Belgium" |
                                             CountryName == "Poland" |
                                             CountryName == "Sweden" |
                                             CountryName == "Ireland" |
                                             CountryName == "Austria" |
                                             CountryName == "Denmark")

### Changing United_Kingdom to United Kingdom
rename <- c(United_Kingdom="United Kingdom", Germany="Germany", Italy="Italy",
            France="France", Spain="Spain", Netherlands="Netherlands", Belgium="Belgium",
            Poland="Poland", Sweden="Sweden", Ireland="Ireland", Austria="Austria",
            Denmark="Denmark")

cases_data_Europe$CountryName <- as.character(rename[cases_data_Europe$CountryName])

### Only selecting the columns that are interesting
cases_data_Europe_small <- cases_data_Europe %>% select(Date, CountryName, notification_rate_per_100000_population_14.days)

## Adding the Trends Data ----
trends_data <- read.csv("GoogleTrends.csv", sep = ";")

### Forcing correct date format
trends_data$Date = as.Date(trends_data$Date, format = "%d.%m.%Y")+1

## Renaming the Sum column to Publicity
names(trends_data)[names(trends_data)=="Sum"] <- "Publicity"

## Only selecting the columns that are interesting
trends_data <- trends_data %>% select(Date, CountryName, Publicity)

## Merging policy mobility trends and cases data on date and country ----

data = policy_data_Europe_small %>% 
  full_join(mobility_data_Europe_small, by = c("Date", "CountryName")) %>%
  full_join(cases_data_Europe_small, by = c("Date", "CountryName")) %>%
  full_join(trends_data, by = c("Date", "CountryName"))

## Adding a column with Compliance-index ----

data$ComplianceIndex <- 100 - (data$StringencyIndex + data$MobilityIndex)

### Rescaling the Compliance-index to assume values between 0 and 100

data$ComplianceIndex <- rescale(data$ComplianceIndex, to = c(0, 100))


```

# Introduction

The COVID-19-Pandemic is an ongoing challenge for policymakers. Especially, because overcoming it is so dependent on the policy-compliance of each and every individual. However, how is the state of this policy-compliance in Europe? Does it differ across different countries and how does it evolve over time?

To find an answer to these questions, the following analysis will assess the following hypotheses.

H1\textsubscript{0}: There **is no** difference in compliance with mobility-reducing policies across European countries.

H1\textsubscript{a}: There **is** a difference in compliance with mobility-reducing policies across European countries.

H2\textsubscript{0}: There **is no** linear trend in such compliance over time.

H2\textsubscript{a}: There **is** a linear trend in such compliance over time.

This analysis combines data from the Google COVID-19 Community Mobility Reports [@COVID19CommunityMobility2020] and data from the Oxford Policy Tracker [@haleOxfordCOVID19Government2020]. The Google Mobility Report charts "movement trends over time by geography, across different categories."[@COVID19CommunityMobility2020] For the sake of this analysis the movement trends of the categories retail and recreation, supermarkets and pharmacies, public transport and workplaces were included and averaged without weighting. The Stringency Index was obtained from the Oxford Policy Tracker and records information on "the strictness of 'lockdown style' policies that primarily restrict people's behaviour."[@haleOxfordCOVID19Government2020]

For this analysis, policy-compliance is approximated as the difference between the level of policy-stringency and the observed reduction in mobility. 

`Compliance = 100 - (Policy Stringency Index - %-change in Mobility)`

To assess our hypothesis more thoroughly, policy-compliance is additionally corrected for COVID-19-cases [@Data14DayNotification2020] and public attention on the topic, as approximated by Google search queries with the terms *"covid"* and *"corona"* [@GoogleTrends].

Countries included in the analysis were chosen based on their membership to the EU (including the United Kingdom) and their GDP.

The following diagram shows how in March, April and May lockdown-style policies were introduced and how mobility was reduced in concordance in almost all the countries assessed. However, beginning immediately from there on, mobility started to increase again, with the gap between policy stringency and observed mobility-reduction widening - signaling reduced levels of policy-compliance.

```{r faceted plot Stringency and Mobility, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
## Create a faceted plot with Stringency and Mobility for European Countries ----

ggplot(data = data) +
  geom_line(aes(x = Date, y = StringencyIndex), color = "black", linetype = "longdash", size = 0.5) +
  geom_smooth(aes(x = Date, y = -1*MobilityIndex), method = "loess", span = 1/10, 
              se = FALSE, size = 0.5, color = "black", linetype = "solid") +
  facet_wrap(~CountryName) +
  xlab("Month") +
  scale_x_date(breaks = date_breaks("3 months"),
               labels = date_format("%b\n%Y")) +
  scale_y_reverse(name = "Policy Stringency Index (dashed line)",
                  sec.axis = sec_axis(~.*-1, name = "Mobility %-change vs Baseline (solid line)")) +
  theme(
    axis.title.x = element_text(vjust = 0.5),
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.y.right = element_text(color = "black", size=12, vjust = 3),
    text=element_text(family="CMU Sans Serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey", linetype = "dotted"),
    strip.background = element_rect(fill="white"),
    plot.title = element_text(size = 11, face = "bold")
  ) +
  geom_hline(yintercept = 0, color = "black", size = 0.5) +
  ggtitle("Development of Policy Stringency and Mobility Reduction in Europe")
```


# Variation in Policy-Compliance

In order to build our model, we can make a preliminary, bivariate analysis that shows the variance of policy compliance across different European countries. 

The boxplot shows us that, indeed, the levels of policy compliance across European countries differ while the intra-group variance is roughly the same.

```{r boxplot compliance, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
## Create a Boxplot of Compliance by Country ----
ggplot(data = data, aes(x = CountryName, y = ComplianceIndex)) +
  geom_boxplot() +
  xlab("Country") +
  ylab("Policy Compliance Index") +
  ggtitle("Policy Compliance in Europe") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        text=element_text(family="CMU Sans Serif"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour = "grey", linetype = "dotted"),
        plot.title = element_text(size = 11, face = "bold"))
```

Additionally, as we want to assess the change of policy compliance across time in Europe, we can fit a linear regression line to the compliance data. This shows us that in all of the European countries analyzed, compliance with mobility-reducing-policies has reduced.

```{r facetted plot compliance, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
## Create a faceted plot with Compliance for European Countries ----
ggplot(data) +
  geom_smooth(aes(x = Date, y = ComplianceIndex), method = "loess", se = FALSE, span = 1/10, 
              color = "black", linetype = "solid", size = 0.2) +
  geom_smooth(aes(x = Date, y = ComplianceIndex), method = "lm", se = FALSE, 
              color = "black", linetype = "solid", size = 1) + 
  facet_wrap(~CountryName) +
  xlab("Month") +
  scale_x_date(breaks = date_breaks("3 months"),
               labels = date_format("%b\n%Y")) +
  scale_y_continuous(name = "Policy Compliance Index") +
  theme(
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.x = element_text(size=12, vjust = 0.5),
    text=element_text(family="CMU Sans Serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey", linetype = "dotted"),
    strip.background = element_rect(fill="white"),
    plot.title = element_text(size = 11, face = "bold")) +
  ggtitle("Policy Compliance in Europe Across Time")
```

# Modelling
## Linear Non-Hierachical Model

The following non-hierarchical model predicts compliance as the dependent variable, with the Stringency Index, Date and Country as the independent variables.

\begin{equation}
{Compliance} = {\beta}_0 + {\beta}_1 * {Date} + {\beta}_2 * {Country} + {\epsilon}
\end{equation}

```{r Non-Hiearchical Model 1, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
## Fit the model
lm_out <- lm(data = data, ComplianceIndex ~ Date + CountryName)

## Look at the results
summary(lm_out)

## Call:
## lm(formula = ComplianceIndex ~ Date + CountryName, data = ## data)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -52.809  -6.508  -0.297   6.226  40.292 
## 
## Coefficients:
##                             Estimate Std. Error t value ## Pr(>|t|)    
## (Intercept)                1.223e+03  3.575e+01  34.211  < ## 2e-16 ***
## Date                      -6.291e-02  1.936e-03 -32.489  < ## 2e-16 ***
## CountryNameBelgium        -3.599e+00  8.334e-01  -4.318 ## 1.61e-05 ***
## CountryNameDenmark        -9.119e+00  8.314e-01 -10.969  < ## 2e-16 ***
## CountryNameFrance         -7.065e+00  8.314e-01  -8.497  < ## 2e-16 ***
## CountryNameGermany        -1.196e+01  8.314e-01 -14.385  < ## 2e-16 ***
## CountryNameIreland        -5.587e+00  8.341e-01  -6.698 ## 2.43e-11 ***
## CountryNameItaly          -7.642e+00  8.348e-01  -9.154  < ## 2e-16 ***
## CountryNameNetherlands    -5.099e+00  8.444e-01  -6.039 ## 1.70e-09 ***
## CountryNamePoland         -6.166e+00  8.327e-01  -7.404 ## 1.63e-13 ***
## CountryNameSpain          -4.647e+00  8.348e-01  -5.566 ## 2.79e-08 ***
## CountryNameSweden         -1.021e+01  8.314e-01 -12.279  < ## 2e-16 ***
## CountryNameUnited Kingdom -2.807e+00  8.314e-01  -3.376 ## 0.000742 ***
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## Residual standard error: 10.32 on 3646 degrees of freedom
##   (670 observations deleted due to missingness)
## Multiple R-squared:  0.2791,	Adjusted R-squared:  0.2768 
## F-statistic: 117.7 on 12 and 3646 DF,  p-value: < 2.2e-16
```

It yields the following result.

| Variable                  | Coefficient | Std. Error | P-Value |
| ------------------------- | ----------- | ---------- | ------- |
| Date                      | -6.291e-02  | 1.936e-03  | \*\*\*  |
| Belgium                   | -3.599e+00  | 8.334e-01  | \*\*\*  |
| Denmark                   | -9.119e+00  | 8.314e-01  | \*\*\*  |
| France                    | -7.065e+00  | 8.314e-01  | \*\*\*  |
| Germany                   | -1.196e+01  | 8.314e-01  | \*\*\*  |
| Ireland                   | -5.587e+00  | 8.341e-01  | \*\*\*  |
| Italy                     | -7.642e+00  | 8.348e-01  | \*\*\*  |
| Netherlands               | -5.099e+00  | 8.444e-01  | \*\*\*  |
| Poland                    | -6.166e+00  | 8.327e-01  | \*\*\*  |
| Spain                     | -4.647e+00  | 8.348e-01  | \*\*\*  |
| Sweden                    | -1.021e+01  | 8.314e-01  | \*\*\*  |
| United Kingdom            | -2.807e+00  | 8.314e-01  | \*\*\*  |
R\textsuperscript{2} = 0.28;
Adj. R\textsuperscript{2} = 0.28;
Num. obs. = 520;
∗∗∗ p < 0.001; ∗∗ p < 0.01; ∗ p < 0.0

The coefficients for Data and the individual Countries, when compared to Austria, are statistically significant. The coefficient for date is negative, with only little standard error. Also the coefficients for the countries are negative, signalling that Austria showed the highest level of compliance, with Germany and Sweden at the bottom of the list. We can control this relationship for the influence of public attention to the pandemic and for COVID-19 cases per 100.000 and 14 days.

\begin{equation}
{Compliance} = {\beta}_0 + {\beta}_1 * {Date} + {\beta}_2 * {Country} + {\beta}_3 * {Public Attention} + {\beta}_4 * {Cases/(100.000 * 14 days)} + {\epsilon}
\end{equation}

```{r Non-Hiearchical Model 2, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
## Fit the model
lm_out <- lm(data = data, ComplianceIndex ~ Date + CountryName + Publicity + notification_rate_per_100000_population_14.days)

## Look at the results
summary(lm_out)

## Call:
## lm(formula = ComplianceIndex ~ Date + CountryName + Publicity + 
##     notification_rate_per_100000_population_14.days, data = data)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -36.552  -6.433  -0.435   5.822  37.089 
## 
## Coefficients:
##                                                   Estimate Std. Error t value Pr(>|t|)    
## (Intercept)                                      1.381e+03  1.155e+02  11.953  < 2e-16 ***
## Date                                            -7.142e-02  6.264e-03 -11.401  < 2e-16 ***
## CountryNameBelgium                              -3.211e+00  2.122e+00  -1.514 0.130771    
## CountryNameDenmark                              -8.285e+00  2.124e+00  -3.902 0.000108 ***
## CountryNameFrance                               -7.145e+00  2.112e+00  -3.383 0.000773 ***
## CountryNameGermany                              -1.106e+01  2.124e+00  -5.209 2.77e-07 ***
## CountryNameIreland                              -4.066e+00  2.133e+00  -1.906 0.057210 .  
## CountryNameItaly                                -8.116e+00  2.165e+00  -3.748 0.000199 ***
## CountryNameNetherlands                          -2.967e+00  2.142e+00  -1.385 0.166606    
## CountryNamePoland                               -5.667e+00  2.169e+00  -2.613 0.009253 ** 
## CountryNameSpain                                -4.790e+00  2.239e+00  -2.139 0.032904 *  
## CountryNameSweden                               -9.533e+00  2.113e+00  -4.511 8.03e-06 ***
## CountryNameUnited Kingdom                       -3.250e+00  2.126e+00  -1.529 0.126923    
## Publicity                                       -1.085e-02  2.393e-02  -0.453 0.650518    
## notification_rate_per_100000_population_14.days  4.710e-03  2.557e-03   1.842 0.066033 .  
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## Residual standard error: 9.907 on 505 degrees of freedom
##   (3809 observations deleted due to missingness)
## Multiple R-squared:  0.3102,	Adjusted R-squared:  0.2911 
## F-statistic: 16.22 on 14 and 505 DF,  p-value: < 2.2e-16

```

It yields the following result. Except for Date, none of the continuous variables Public Attention and Cases per 100,000 and 14 Days are statistically significant. Date is statistically significant at the 0.05 level. Its coefficient shows that when holding Publicity, Cases and Country constant, Compliance decreases on average by 0.07 points each day. Interestingly, some of the coefficients for individual countries are not significant anymore when controlling for public attention and COVID-19 cases. In the following, a random intercept will be included for country, in order to improve model fit.

 | Variable                      | Coefficient | Std. Error | P-Value |
 | ----------------------------- | ----------- | ---------- | ------- |
 | Date                          | -7.142e-02  | 6.264e-03  | \*\*\*  |
 | Belgium                       | -3.211e+00  | 2.122e+00  |         |
 | Denmark                       | -8.285e+00  | 2.124e+00  | \*\*\*  |
 | France                        | -7.145e+00  | 2.112e+00  | \*\*\*  |
 | Germany                       | -1.106e+01  | 2.124e+00  | \*\*\*  |
 | Ireland                       | -4.066e+00  | 2.133e+00  | .       |
 | Italy                         | -8.116e+00  | 2.165e+00  | \*\*\*  |
 | Netherlands                   | -2.967e+00  | 2.142e+00  |         |
 | Poland                        | -5.667e+00  | 2.169e+00  | \*\*    |
 | Spain                         | -4.790e+00  | 2.239e+00  | \*      |
 | Sweden                        | -9.533e+00  | 2.113e+00  | \*\*\*  |
 | United Kingdom                | -3.250e+00  | 2.126e+00  |         |
 | Public Attention              | -1.085e-02  | 2.393e-02  |         |
 | Cases per 100,000 and 14 Days | 4.710e-03   | 2.557e-03  | .       |
R\textsuperscript{2} = 0.31;
Adj. R\textsuperscript{2} = 0.29;
Num. obs. = 520;
∗∗∗ p < 0.001; ∗∗ p < 0.01; ∗ p < 0.05

## Linear Multi-Level Model

As we can see in the regression output of our non-hierarchical model, the fixed effects estimated for the different expressions of our country variable show substantial differences, with Germany showing more than 10 points less policy-compliance than our reference country Austria.

To better assess the trends of policy-compliance in Europe over time, we can create a hierarchical model, allowing both random intercepts for the subgroups of Country as our grouping variable and random slopes along the Date variable.

\begin{equation}
{Compliance}_{ij} = {\beta}_0 + {\beta}_1 * {Date}_{ij} + {\beta}_2 * {Public Attention}_{ij} + {\beta}_3 * {Cases/(100,000 * 14 days)}_{ij} + {\beta}_{1j} * {Date} + {Country}_{0j} + {\epsilon}_{0ij}
\end{equation}

```{r Hierarchical model, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
## How does the Compliance change over time in Europe? This model has not enough independent variables ----

lmer_out <- lmer(data = data, ComplianceIndex ~ Date + Publicity + notification_rate_per_100000_population_14.days +  (Date | CountryName))

### Looking at the results
summary(lmer_out)

## Linear mixed model fit by REML ['lmerMod']
## Formula: ComplianceIndex ~ Date + Publicity + notification_rate_per_100000_population_14.days +  
##     (Date | CountryName)
##    Data: data
## 
## REML criterion at convergence: 3908.6
## 
## Scaled residuals: 
##     Min      1Q  Median      3Q     Max 
## -3.7395 -0.6504 -0.0370  0.5971  3.7889 
## 
## Random effects:
##  Groups      Name        Variance  Std.Dev.  Corr 
##  CountryName (Intercept) 9.644e+01 9.8205986      
##              Date        9.074e-07 0.0009526 -1.00
##  Residual                9.643e+01 9.8196921      
## Number of obs: 520, groups:  CountryName, 12
## 
## Fixed effects:
##                                                   Estimate Std. Error t value
## (Intercept)                                      1.376e+03  1.145e+02  12.015
## Date                                            -7.148e-02  6.213e-03 -11.505
## Publicity                                       -1.027e-02  2.362e-02  -0.435
## notification_rate_per_100000_population_14.days  4.744e-03  2.533e-03   1.873
## 
## Correlation of Fixed Effects:
##             (Intr) Date   Pblcty
## Date        -1.000              
## Publicity   -0.155  0.148       
## n___100000_  0.610 -0.610 -0.288
## optimizer (nloptwrap) convergence code: 0 (OK)
## boundary (singular) fit: see ?isSingular

### Getting p-values

2*pt(11.505, 520, lower.tail = F)

2*pt(0.435, 520, lower.tail = F)

2*pt(1.873, 520, lower.tail = F)

## > 2*pt(11.505, 520, lower.tail = F)
## [1] 1.906225e-27
## > 2*pt(0.435, 520, lower.tail = F)
## [1] 0.6637429
## > 2*pt(1.873, 520, lower.tail = F)
## [1] 0.06162918

```


| Random effects | -         |        - |         - |
| -------------- | --------- | --------:| ---------:|
| Groups         | Name      | Variance | Std. Dev. |
| Country        | Intercept |    96.44 |      9.82 |
|                | Date      |     0.00 |      0.00 |
| Residual       |           |    96.43 |      9.82 |
Number of observations: 520, Number of groups: 12

| Fixed effects                 |        - |          - |       - |    -    |
| ------------------------------| --------:| ----------:| -------:|:-------:|
|                               | Estimate | Std. Error | t-value | p-value |
| Date                          |    -0.07 |       0.01 |  -11.51 |   ∗∗∗   |
| Public Attention              |    -0.01 |       0.02 |   -0.44 |         |
| Cases per 100,000 and 14 Days |     0.00 |       0.00 |    1.87 |         |
∗∗∗ p < 0.001; ∗∗ p < 0.01; ∗ p < 0.05

As we can see, the coefficients of the fixed effects at level one have not changed substantially. Still, the fixed effect of Date is the only one that is statistically significant.

The following plot shows the rate at which compliance changes per day in different European countries, according to our hierarchical model, that controls for public attention and COVID-19-cases.

```{r slope of compliance-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}

# Model
lmer_out <- lmer(data = data, ComplianceIndex ~ Date + Publicity + notification_rate_per_100000_population_14.days +  (Date | CountryName))

## Extracting out the fixed-effect slope for Date
Date_slope <- fixef(lmer_out)['Date']

## Extracting out the random-effect slopes for Country
Country_slope <- ranef(lmer_out)$CountryName

## Creating a new column for the slope
Country_slope$slope <- Country_slope$Date + Date_slope

## Using row names to create a county name column
Country_slope$country <- rownames(Country_slope)

## Ordered county-level factor based upon slope values
Country_slope$country_plot <- factor(Country_slope$country, 
                                   levels = Country_slope$country[order(Country_slope$slope)])

## Plotting the results
ggplot(data = Country_slope, aes(x = country_plot, y = slope)) + 
  geom_point() +
  coord_flip() + 
  theme_bw() + 
  ylab("Change in Compliance per Day")  +
  xlab("Country") +
  theme(
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.x = element_text(size=12, vjust = 0.5),
    text=element_text(family="CMU Sans Serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey", linetype = "dotted"),
    plot.title = element_text(size = 11, face = "bold")) +
  ggtitle("Rate of Change in Compliance in Europe")
```


## Model comparison

```{r Likelihood ratio test, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

lrtest(lm_out, lmer_out)

## Likelihood ratio test
## 
## Model 1: ComplianceIndex ~ Date + CountryName + Publicity + notification_rate_per_100000_population_14.days
## Model 2: ComplianceIndex ~ Date + Publicity + notification_rate_per_100000_population_14.days + 
##     (Date | CountryName)
##   #Df  LogLik Df Chisq Pr(>Chisq)    
## 1  16 -1922.8                        
## 2   8 -1954.3 -8 63.15  1.119e-10 ***
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

Finally, to compare our hierarchical model-fit with the model-fit of our non-hierarchical model, we can compute a likelihood ratio test. This yields the following result.

`Model 1: ComplianceIndex ~ Date + Country + Public Attention + Cases per 100,000 and 14 Days`

`Model 2: ComplianceIndex ~ Date + Public Attention + Cases per 100,000 and 14 Days + (Date | Country)`

| Model | Df  | LogLik  | Chisq  | P-Value       |
| ----- | --- | ------- | ------ | --------------|
| 1     | 16  | -1922.8 |        |               |
| 2     | 8   | -1954.3 | 63.15  | 1.119e-10 *** |
∗∗∗ p < 0.001; ∗∗ p < 0.01; ∗ p < 0.05

As we can see, the improvement of our hierarchical model over our non-hierarchical model is statistically significant at the 0.05 level.

# Conclusion

This analysis has shown that, interestingly, levels of compliance with mobility-reducing COVID-19-policies differ across selected European countries and that policy-compliance expresses an alarming, negative trend. Additionally, we were able to show how the *rate* at which policy-compliance decreases differs across European countries.

# References


